---
title: "Demonstration of using GRASS from R"
author: "Floris Vanderhaeghe"
date: "2021-06-17"
site: bookdown::bookdown_site
output: 
  bookdown::html_document2:
    number_sections: yes
    fig_caption: yes
    df_print: paged
    toc: TRUE
    toc_depth: 4
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    includes:
        in_header: header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
execshell <- function(commandstring, intern = FALSE, ...) {
  if (.Platform$OS.type == "windows") {
    res <- shell(commandstring, intern = TRUE, ...)
  } else {
    res <- system(commandstring, intern = TRUE, ...)
  }
  if (!intern) {
    if (length(res) > 0) cat(res, sep = "\n") else return(invisible())
    } else return(res)
}
```

The source code of this file can be found [at GitHub](https://github.com/florisvdh/demos/tree/master/src/20210617_grass).

An accompanying presentation is at <https://florisvdh.github.io/presentations/20210617_grass>.

```{r message=FALSE}
library(rgrass7)
use_sf()
library(tmap)
library(dplyr)
```

# Initializing a temporary GRASS project

```{r}
gisbase_grass <- ifelse(.Platform$OS.type == "windows", 
                        link2GI::paramGRASSw()$gisbase_GRASS[1],
                        link2GI::paramGRASSx()$gisbase_GRASS[1])
initGRASS(gisBase = gisbase_grass,
          home = tempdir(), 
          mapset = "PERMANENT")
stringexecGRASS("g.proj -c --quiet epsg=31370")
```

# GRASS workflow

## Loading data from WFS

In case GRASS will work directly on pre-existing data in the cloud or on disk,
it's more efficient to let GRASS load those directly than to read them in R and
then use `rgrass7::writeXXX()`.
The latter will first write a temporary GPKG or GeoTIFF file from R, after which
GRASS is instructed to read that file.
So that route is most convenient for intermediate data created in R.

```{r}
execshell("v.in.ogr input='https://geoservices.informatievlaanderen.be/overdrachtdiensten/BWK/wfs?service=WFS&request=GetFeature&typename=BWK%3ABwkhab&bbox=130000%2C184000%2C138000%2C194000' output=habbox -o")
# execGRASS() currently has bug because of '&' in URL:
# url <- "https://geoservices.informatievlaanderen.be/overdrachtdiensten/BWK/wfs?service=WFS&request=GetFeature&typename=BWK%3ABwkhab&bbox=130000%2C184000%2C138000%2C194000"
# execGRASS("v.in.ogr",
#           "o",
#           input = url,
#           output = "habbox")
```


```{r}
stringexecGRASS("v.info map=habbox")
stringexecGRASS("v.info -c map=habbox")
```

## Dissolving polygon borders

```{r warning=FALSE}
stringexecGRASS("v.dissolve input=habbox output=habeval column=EVAL")
stringexecGRASS("v.db.select map=habeval")
```

## Rasterizing result

```{r}
# note that stringexecGRASS() cannot split '-sa' as 
# '-s -a', which is needed for execGRASS()
execshell("g.region vector=habeval res=1 -sa
           g.region -p
           v.to.rast input=habeval output=reval use=cat label_column=EVAL memory=2048")
```

# Read results as R object

```{r warning=FALSE}
result <- readVECT("habeval")
```

```{r fig.cap="Result after dissolve step (dataset `habeval`)"}
tm_shape(result) + tm_polygons("EVAL")
```


# Used environment

## Platform

```{r session-info, results = "asis", echo=FALSE}
si <- sessioninfo::session_info()
p <- si$platform %>%
  do.call(what = "c")
if ("sf" %in% si$packages$package) {
  p <- c(p, sf::sf_extSoftVersion())
  names(p)[names(p) == "proj.4"] <- "PROJ"
}
if ("rgrass7" %in% si$packages$package) {
  p <- c(p, GRASS = link2GI::findGRASS()[1, "version"])
}
sprintf("- **%s**: %s\n", names(p), p) %>%
  cat(sep = "")
```

## R packages

```{r results = "asis", echo=FALSE}
si$packages %>%
    as_tibble %>%
    select(package, loadedversion, date, source) %>%
pander::pandoc.table(caption = "(\\#tab:sessioninfo)Loaded R packages",
                     split.table = Inf)
```
